### 2018-11-15 lab_notebook.Rmd
creating turf/itq data set almost done

---
title: "lab_notebook"
author: "Juliette Verstaen"
date: "November 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading packages
```{r}

suppressPackageStartupMessages({
  library(tidyverse)
  library(tidyr)
  library(here)
  library(ggplot2)
  library(dplyr)
  library(readxl)
  library(devtools)
  library(plotly)
  library(base)
  library(rio)
  library(plyr)
})

```

#converting projection csv into rds
no longer need the code because have already used it to create the files
```{r}

#Upside data
#projection_data_csv <- read_csv("data/ProjectionData.csv")

#saveRDS(projection_data_csv, "projection_data.rds")

#WDPA_Nov2018.csv <- read_csv("data/WDPA_Nov2018.csv")

#saveRDS(WDPA_Nov2018.csv, "WDPA_Nov2018.rds")

```

#Load Data
```{r}

#Corbet data: when countries began using RBFs
countries_ITQ <- read_excel("data/Countries_ITQ.xlsx")

#Upside data
projection_data <- readRDS("data/projection_data.rds")

#MPA data
MPA <- readRDS("data/WDPA_Nov2018.rds")

#RAMS
load("data/DBdata-model_fits_included.RData")

#number_species <- unique(stock$scientificname)
#341 different species

#cooperative data from "Conservation incentives and collective choices in cooperative fisheries" paper (Ovando et al 2013)
jc_coop <- read_csv("data/clean_cooperatives_data.csv")

#EDF turf data
edf_turf <- read_excel("data/EDF_TURF_data.xlsx")

```

Important/Necessary Data
tcbest.data #the best estimate of catch for that stock
tbbest.dat # the best estimate of biomass for that stock
divbpref # B/bmsy
divupref.data # u/umsy

Juliette Notes:

Handy functions for updating data frames:
match()
separate()

Useful Websites:
https://uc-r.github.io/tidyr

#RAM data wrangling
```{r}
ram <- timeseries_values_views %>%
  select("stocklong", "year", "TBbest", "TCbest", "BdivBmsypref", "UdivUmsypref") %>%
  add_column("RAM")

colnames(ram) <- c( "CommName", "Year", "Biomass", "Catch", "BvBmsy", "FvFmsy", "DBase")

#ram_2012 <- ram %>%
#  filter(Year > "2012") 
##this wont work because projection data has projected out past 2012. need to UPDATE

#updated 
#projection_ram <- bind_rows(list(projection_data, ram_2012))

2012-1950
#62
2050- 2012
#38
162+38

acheck <- filter(projection_ram, Year == "2013") %>%
  select("IdOrig", "CommName", "Year", "Biomass")

acheck_count <- unique(acheck$IdOrig)
acheck_count
#4714

yearcheck <- unique(projection_data$Year)
yearcheck
#101 years starts 1950 end 2050

join_project_ram <- join(projection_data, ram, by = c("CommName", "Year"), type = "full")

projectramcheck <- filter(join_project_ram, IdOrig == "NA")
#returns no rows with NA, could still be right if updated projected

#Checking to see if better way to do join
# end up with WAY to many rows, something is off
#unite_projection_commname <- projection_data %>% 
#  unite(name_year, CommName, Year, sep = "-")
#unite_ram_commname <- ram %>% 
#  unite (name_year, CommName, Year, sep = "-")
#join_project_ram_2 <- join(unite_projection_commname, unite_projection_commname, by = "name_year", type = "left")

match_project_ram <- match(projection_data, ram)

```

#Data Wrangling: Projection and Countries ITQ
```{r}

#seperate out IdOrig column for Projection Data
separate_projection <- projection_ram %>% separate(IdOrig, c("agency", "stock", "t0", "tf", "data_enterer"))
    ###unsure what the last column actually is, double check

head(separate_projection, 10)

#reunite just the two parts we care about: agency and stock
unite_projection <- separate_projection %>% unite(assess_id_short, agency, stock, sep = "-")

#simplifying Countries ITQ with unique columns to prepare to merge
countries_ITQ_simple <- select(countries_ITQ, "assess_id_short", "year", "yearitq", "region_id", "idnumber", "harvest_fb", "itq_now", "iq", "itq", "ivq", "coop", "communityquota", "turf", "individualeffort", "iteq", "totalappliedeffort", "producerorg", "aquaculture", "programstart")

#rename year to be capitalized to merge with projection data
colnames(countries_ITQ_simple) <- c("assess_id_short", "Year", "yearitq", "region_id", "idnumber", "harvest_fb", "itq_now", "iq", "itq", "ivq", "coop", "communityquota", "turf", "individualeffort", "iteq", "totalappliedeffort", "producerorg", "aquaculture", "programstart")

ITQ_projection_merge <- merge(unite_projection, countries_ITQ_simple, by = c("assess_id_short", "Year"), all = TRUE)

#cleaning it up
ITQ_projection <- ITQ_projection_merge %>% unite(IdOrig, assess_id_short, t0, tf, data_enterer, sep = "-")

```

Playing with Data

#Total number of species with stocks
#Number of stocks with ITQs implemented
```{r}
IdOrig <- unique(ITQ_projection$SciName)
IdOrig
#1228 species total

ITQ_projection_implemented <- filter(ITQ_projection, programstart != "NA")

IdOrig_implemented <- unique(ITQ_projection_implemented$SciName)
IdOrig_implemented
#80 species with ITQ implemented
```

#Graphs for fun

ABARES_BGRDRSE
implemented ITQs in 1992
```{r}

ABARES_BGRDRSE <- filter(ITQ_projection, IdOrig == "ABARES-BGRDRSE-1960-2011-CHING")

graph_ABARES_BGRDRSE <-ggplot(ABARES_BGRDRSE, aes(x=Year, y=Biomass))+
  geom_line(size=1)+
  ggtitle("ABARES_BGRDRSE (Australia")+
  geom_point()+
  geom_line()+
  labs(x="Year", y="Yearly total biomass")+
  theme_classic()+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand=c(0,0))+
  geom_vline(xintercept = 1992)+
  theme(legend.title=element_blank())

ggplotly(graph_ABARES_BGRDRSE)

#figure out an if then statement for line?

```

ADFG-HERRTOGE
never implemented ITQs
```{r}
ADFG_HERRTOGE <- filter(ITQ_projection, IdOrig == "ADFG-HERRTOG-1990-2010-HIVELY")

graph_ADFG_HERRTOGE <-ggplot(ADFG_HERRTOGE, aes(x=Year, y=Biomass))+
  geom_line(size=1)+
  ggtitle("ADFG-HERRTOGE (USA)")+
  geom_point()+
  geom_line()+
  labs(x="Year", y="Yearly total biomass")+
  theme_classic()+
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand=c(0,0))+
  theme(legend.title=element_blank())

ggplotly(graph_ADFG_HERRTOGE)

#figure out an if then statement for line?
```

Important/Necessary Data
tcbest.data #the best estimate of catch for that stock
tbbest.dat # the best estimate of biomass for that stock
divbpref # B/bmsy
divupref.data # u/umsy

Attempts at updating ITQ_projection with Rams
#asked ana about this -> waiting to hear back
```{r}

##this doesn't work; gives you vector not updated data frame
RAMS_ITQ_projection <- match(ITQ_projection_year$CommYear, ram_merge$CommYear)

#what happens when i just merge the two?
#test_merge <- merge(ITQ_projection_year, ram_merge, by = c("Name_Year"))

```

##edf turf data wrangling
# waiting for data from Gina if its more comprehensive than the EDF one
```{r}

edf_turf_unite <- edf_turf %>% unite(SciName, genus, species, sep = " ")

edf_turf_simple <- edf_turf_unite %>%
  add_column("turf") %>%
  select("country", "start_date", "SciName") %>%
  gsub('\\bNA NA\\b', 'NA')

colnames(edf_turf_simple) <- c("Country", "turfstart", "SciName")

##want to merge the edf simple with ITQ/Projections but maybe wait for Gina to respond with her TURF data
##in ITQ data says only one species of fish in japan has TURFs but not true

edf_projection_itq <- merge(ITQ_projection, edf_turf_simple, by = c("Country", "SciName"))
#only three species from the EDF data match with the species from the upside/projection data; all in Japan

```

Need to create new dataset (half upside half ITQ): Species, turf/itq, country,and then matching all the parameters with the species

################
################
###############
PART 2

From updating_projection.R script file
```{r}

#Load Data
#Upside data
projection_data <- readRDS("data/projection_data.rds")

#RAMS
load("data/DBdata-model_fits_included.RData")

#Corbet data: when countries began using RBFs
countries_ITQ <- read_excel("data/Countries_ITQ.xlsx")

#RAM Data Tiddying
ram <- timeseries_values_views %>%
  select("stocklong", "year", "TBbest", "TCbest", "BdivBmsypref", "UdivUmsypref") %>%
  add_column("RAM")

colnames(ram) <- c( "CommName", "Year", "Biomass", "Catch", "BvBmsy", "FvFmsy", "DBase")

#Update Projection with RAM
projection_updated <- join(projection_data, ram, by = c("CommName", "Year"), type = "full")

#Merge Projection and ITQ
#seperate out IdOrig column for Projection Data
separate_projection <- projection_updated %>% separate(IdOrig, c("agency", "stock", "t0", "tf", "data_enterer"))

head(separate_projection, 10)

#reunite just the two parts we care about: agency and stock
unite_projection <- separate_projection %>% unite(assess_id_short, agency, stock, sep = "-")

#simplifying Countries ITQ with unique columns to prepare to merge
countries_ITQ_simple <- select(countries_ITQ, "assess_id_short", "year", "yearitq", "region_id", "idnumber", "harvest_fb", "itq_now", "iq", "itq", "ivq", "coop", "communityquota", "turf", "individualeffort", "iteq", "totalappliedeffort", "producerorg", "aquaculture", "programstart")

#rename year to be capitalized to merge with projection data
colnames(countries_ITQ_simple) <- c("assess_id_short", "Year", "yearitq", "region_id", "idnumber", "harvest_fb", "itq_now", "iq", "itq", "ivq", "coop", "communityquota", "turf", "individualeffort", "iteq", "totalappliedeffort", "producerorg", "aquaculture", "programstart")

ITQ_projection_merge <- merge(unite_projection, countries_ITQ_simple, by = c("assess_id_short", "Year"), all = TRUE)

#cleaning it up
ITQ_projection <- ITQ_projection_merge %>% unite(IdOrig, assess_id_short, t0, tf, data_enterer, sep = "-")

#simplified projection data
ITQ_projection_simple <- ITQ_projection %>%
  select("Country", "Year", "CommName", "SciName", "SpeciesCat", "CatchShare", "Catch", "Biomass", "BvBmsy", "FvFmsy","MSY", "Price", "g", "k", "c", "phi", "yearitq", "itq_now","iq", "itq", "ivq", "turf", "programstart")

#create RDS from dataframe
#saveRDS(ITQ_projection_simple, "projection_updated.rds")

#upside updated
upside_2018_updated <- ITQ_projection %>%
  select("Country", "Year", "CommName", "SciName", "SpeciesCat", "CatchShare", "Catch", "Biomass", "BvBmsy", "FvFmsy","MSY", "Price", "g", "k", "c", "phi")

#create RDS from dataframe
#saveRDS(upside_2018_updated, "upside_2018_updated")

```

1) pull the ITQ, IVQ, IQ fisheries out of upside.
2) pull the TURF fisheries from EDF.
3) For each fishery, we want species, location, and ITQ vs. TURF.

```{r}

# create dataframe with 
itq <- ITQ_projection %>%
  select("Country", "SciName", "itq_now","iq", "itq", "ivq", "turf", "programstart")

#Load EDF TURF data

edf_turf <- read_excel("data/EDF_TURF_data.xlsx")

#load in DiscoverTURFs data

d_turfs <- read_excel("data/DiscoverTURFs_simple.xlsx")

```

```{r}

edf_turf_simple <- select(edf_turf, "country", "start_date", "SciName", "turf")
#create new column that says "turf" and TRUE or FALSE

colnames(edf_turf_simple) <- c("Country", "programstart", "SciName", "turf")

#combining edf turf data from upside data
edf_projection_1 <- join(edf_turf_simple, itq, by = c("Country", "programstart", "SciName", "turf"), type = "full")

#combine edf/upside with discover turfs
all_turf_1 <- join(edf_projection_1, d_turfs, by = c("Country", "SciName", "turf"), type = "full")

#change "blank" NAs to real NAs
all_turf_1[is.na(all_turf_1)] <- "NA"

#rearrange columns
all_turf_2 <- all_turf_1[, c(1,3,2,5,6,7,8,4)]

#turning NAs from no ITQ data to FALSE
#all_turf_2$itq_now[all_turf_2$itq_now == 0] <- NA

#all_turf_2 %>% replace(is.na(.), FALSE)

#doublecheckworked <- all_turf_2 %>%
#  filter( itq_now == "1" & turf == "TRUE")

```