### 2018-12-09 draft_figures.Rmd
probability functions for itq and turfs

---
title: "Draft_Figures"
author: "Juliette Verstaen"
date: "12/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

suppressPackageStartupMessages({
  library(tidyverse)
  library(tidyr)
  library(here)
  library(ggplot2)
  library(dplyr)
  library(readxl)
  library(devtools)
  library(plotly)
  library(base)
  library(rio)
  library(plyr)
})
```

##From upside data: ITQ versus non ITQ and the size of dot is the catch

Note: Corbett's ITQ data is only until 2012. Chris's upside data is now updated until 2016 with the more recent RAM data that Dan had. I ran all the figure and regressions under two "scenerios":

    1. until 2012
    2. past 2012. extracted information about which fisheries had ITQs in 2012 and applied it   to the same fisheries into the future. This doesn't take into account other fisheries that   get ITQs in place during that time, or if fisheries stop ITQs

"KOBE" plots with ITQs not updated past 2012
```{r}

ITQ_projection <- readRDS("data/ITQ_projection.rds")

f_b_itq <- ITQ_projection %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(rightsbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

F_B_graph <- ggplot(data = f_b_itq, aes( x=BvBmsy, y=FvFmsy, colour= rightsbased, size = Catch ))+
  geom_point()+
  labs(x = "B/Bmsy", y= "F/Fsmy") +
  theme_minimal()+
  theme(legend.title=element_blank())+
  ylim(-1, 5)+
  xlim(-.3, 3)+
  geom_hline(aes(yintercept=1))+
  geom_vline(aes(xintercept=1))

#F_B_graph
ggplotly(F_B_graph)

```

"KOBE" plots with ITQs updated past 2012
```{r}

ITQ_projection_updated <- readRDS("data/ITQ_projection_updated.rds")

f_b_itq_updated <- ITQ_projection_updated %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(rightsbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

F_B_graph_updated <- ggplot(data = f_b_itq_updated, aes( x=BvBmsy, y=FvFmsy, colour= rightsbased, size = Catch ))+
  geom_point()+
  labs(x = "B/Bmsy", y= "F/Fsmy") +
  theme_minimal()+
  theme(legend.title=element_blank())+
  ylim(-1, 5)+
  xlim(-.3, 3)+
  geom_hline(aes(yintercept=1))+
  geom_vline(aes(xintercept=1))

#F_B_graph
ggplotly(F_B_graph_updated)
```

##From upside: the catch in ITQ versus non ITQ for most recent year

Data until 2012
```{r}
ITQ_projection <- readRDS("data/ITQ_projection.rds")

f_b_itq_countries <- ITQ_projection %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch", "Year", "Country", "IdOrig") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(rightsbased = case_when(
    turf == TRUE ~ "2",
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

countries_itq_2012 <- f_b_itq_countries %>%
  select (Country, rightsbased, Year, IdOrig, Catch) %>%
  filter(Year == "2012")

itq_2012_rightsbasedTRUE <- filter(countries_itq_2012, rightsbased == "1")

itq_2012_rightsbasedFALSE <- filter(countries_itq_2012, rightsbased == "0")

itq_2012_rightsbasedTURF <- filter(countries_itq_2012, rightsbased == "2")

sum(itq_2012_rightsbasedTRUE$Catch, na.rm = TRUE)
#7,420,003

sum(itq_2012_rightsbasedFALSE$Catch, na.rm = TRUE)
#142,231

sum(itq_2012_rightsbasedTURF$Catch, na.rm = TRUE)
#1,301,000
```
2012 total catch:
ITQs: 7,420,003
No ITQ: 142,231
TURF:1,301,000

Data past 2012
```{r}

ITQ_projection <- readRDS("data/ITQ_projection.rds")

f_b_itq_countries <- ITQ_projection %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch", "Year", "Country", "IdOrig") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(rightsbased = case_when(
    turf == TRUE ~ "2",
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

countries_itq_2012 <- f_b_itq_countries %>%
  select (Country, rightsbased, Year, IdOrig, Catch) %>%
  filter(Year == "2012")

ITQ_projection_2016 <- ITQ_projection %>%
  filter(Year == "2016") %>%
  select(IdOrig, Country)

itq_2016_updated <- join(ITQ_projection_2016, countries_itq_2012, by= c("IdOrig", "Country"), type= "full")

itq_2016_rightsbasedTRUE <- filter(itq_2016_updated, rightsbased == "1")

itq_2016_rightsbasedFALSE <- filter(itq_2016_updated, rightsbased == "0")

itq_2016_rightsbasedTURF <- filter(itq_2016_updated, rightsbased == "2")

sum(itq_2016_rightsbasedTRUE$Catch, na.rm = TRUE)
#27,830,933

sum(itq_2016_rightsbasedFALSE$Catch, na.rm = TRUE)
#266,328,075

sum(itq_2016_rightsbasedTURF$Catch, na.rm = TRUE)
#1505000
```

2016 total catch:
ITQs: 27,830,933
No ITQ: 266,328,075
TURF:1,505,000

##Probability of ITQ

```{r}

species_rightsbased <- read_csv("data/species_rightsbased.csv")

gdp_all <- read_csv("/GitHub/global_fisheries_managment/data/gpd.csv")

gdp <- filter(gdp_all, current_gdp != "NA")

species_rightsbased_mutate <- species_rightsbased %>%
  mutate(i_right = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0")) %>%
  mutate(TURF = case_when(
    turf == TRUE ~ "1", 
    turf = FALSE ~"0"
  ))

##why does this value have NAs in the first place? when mutate is does line 1 first and not registering the false = 0 one. but i guess it doesn't matter beacuse its either one or the other? this is not happening for i_right so thats weird

species_rightsbased_mutate$TURF[is.na(species_rightsbased_mutate$TURF)] <- "0"

gdp <- filter(gdp_all, current_gdp != "NA")

join_gdp_rightsbased <- join(gdp, species_rightsbased_mutate, by = c("Country"), type = "full")

gdp_rightsbased <- filter(join_gdp_rightsbased, SpeciesCat != "NA" )

gdp_rightsbased$SpeciesCat <- factor(gdp_rightsbased$SpeciesCat)
gdp_rightsbased$i_right <- factor(gdp_rightsbased$i_right)

itq_glm <- glm(formula = i_right ~ current_gdp + SpeciesCat, family = "binomial", data = gdp_rightsbased)
itq_glm

summary(itq_glm)

```

### this code chunk and the one above should be the same but they are not. figure out why sinc eth two data frames are pulled from the same original info
```{r}

#ISSCAAP_species <- read_csv("data/ISSCAAP_species.csv")
#ITQ_projection <- readRDS("data/ITQ_projection.rds")

#ITQ_projection_isscaap_1 <- join(ISSCAAP_species,ITQ_projection, by = c("SciName", "SpeciesCat"), type = "full" )

#ITQ_projection_isscaap <- ITQ_projection_isscaap_1 %>%
 # filter(Country != "NA")

species_rightsbased <- read_csv("data/species_rightsbased.csv")

gdp_all <- read_csv("/GitHub/global_fisheries_managment/data/gpd.csv")

gdp <- filter(gdp_all, current_gdp != "NA")

species_rightsbased <- ITQ_projection_isscaap %>%
  select("SciName", "Country", "SpeciesCat", "itq", "iq", "ivq", "turf", "Catch") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(i_right = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0")) %>%
  mutate(turf = case_when(
    turf == TRUE ~ "1", 
    turf = FALSE ~"0"
  )) %>%
  select("SciName", "Country", "SpeciesCat", "i_right", "turf", "Catch")

species_rightsbased$turf[is.na(species_rightsbased$turf)] <- "0"

join_gdp_rightsbased <- join(gdp, species_rightsbased, by = c("Country"), type = "full")

gdp_rightsbased <- filter(join_gdp_rightsbased, SpeciesCat != "NA" )

gdp_rightsbased$SpeciesCat <- factor(gdp_rightsbased$SpeciesCat)
gdp_rightsbased$i_right <- factor(gdp_rightsbased$i_right)

itq_glm <- glm(formula = i_right ~ current_gdp + SpeciesCat, family = "binomial", data = gdp_rightsbased)
itq_glm

summary(itq_glm)

```

##Run Turf versus ITQ probably: prob(ITQ). Turf = 1 and ITQ = 0 with the data set that was the turf/itq only one

```{r}
species_rightsbased <- read_csv("data/species_rightsbased.csv")

gdp_all <- read_csv("/GitHub/global_fisheries_managment/data/gpd.csv")

gdp <- filter(gdp_all, current_gdp != "NA")

species_rightsbased_mutate_2 <- species_rightsbased %>%
  mutate(rightbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "0",
    turf == TRUE ~ "1"))

gdp <- filter(gdp_all, current_gdp != "NA")

join_gdp_rightsbased_2 <- join(gdp, species_rightsbased_mutate_2, by = c("Country"), type = "full")

gdp_rightsbased_2 <- filter(join_gdp_rightsbased_2, SpeciesCat != "NA" )

gdp_rightsbased_2$SpeciesCat <- factor(gdp_rightsbased_2$SpeciesCat)
gdp_rightsbased_2$rightbased <- factor(gdp_rightsbased_2$rightbased)

rightbased_glm <- glm(formula = rightbased ~ current_gdp + SpeciesCat, family = "binomial", data = gdp_rightsbased_2)
rightbased_glm

summary(rightbased_glm)

```