### 2018-12-11 draft_figures.Rmd
"kobe" plots 2012 and 2016 total catch 2012 and 2016

---
title: "Draft_Figures"
author: "Juliette Verstaen"
date: "12/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

suppressPackageStartupMessages({
  library(tidyverse)
  library(tidyr)
  library(here)
  library(ggplot2)
  library(dplyr)
  library(readxl)
  library(devtools)
  library(plotly)
  library(base)
  library(rio)
  library(plyr)
})
```

##From upside data: ITQ versus non ITQ and the size of dot is the catch

Notes:
- Corbett's ITQ data is only until 2012. Chris's upside data is now updated until 2016 with the more recent RAM data that Dan had. I ran all the figure and regressions under two "scenerios":

    1. until 2012
    2. past 2012. extracted information about which fisheries had ITQs in 2012 and applied it to the same fisheries into the future. This doesn't take into account other fisheries implement ITQs past 2012, or if fisheries stop ITQs. Only countries with ITQs in 2012 are : Australia, USA, Canda, Japan, Peru, Argentina, South Africa, New Zealand

- for all ITQ related analyzes, limited to fisheries that Corbette had data on. I remove the fisheries that have no information on presence or absence of ITQs. I am unsure if this is biasing the data; should we assume that the fisheries with no information on ITQs have no ITQs or not? This seems like it woud be biased as well. 
    - Corbette's ITQ data includes 375 different fisheries years 1950-2012 for a total of ~15,000 lines of data

"KOBE" plots with ITQs not updated past 2012
All fisheries 1950-2012 
```{r}

ITQ_projection <- readRDS("data/ITQ_projection.rds")

f_b_itq <- ITQ_projection %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(rightsbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

f_b_itq$rightsbased[f_b_itq$rightsbased == "0"]<- "No ITQ"
f_b_itq$rightsbased[f_b_itq$rightsbased == "1"]<- "ITQ"

F_B_graph <- ggplot(data = f_b_itq, aes( x=BvBmsy, y=FvFmsy, colour= rightsbased, size = Catch ))+
  geom_point()+
  labs(x = "B/Bmsy", y= "F/Fsmy") +
  theme_minimal()+
  theme(legend.title=element_blank())+
  ylim(-1, 5)+
  xlim(-.3, 3)+
  geom_hline(aes(yintercept=1))+
  geom_vline(aes(xintercept=1))+
  ggtitle("Fisheries 1950-2012")

#F_B_graph
ggplotly(F_B_graph)

```

"KOBE" plots with ITQs updated past 2012
```{r}

ITQ_projection_updated <- readRDS("data/ITQ_projection_updated.rds")

f_b_itq_updated <- ITQ_projection_updated %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(rightsbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

f_b_itq_updated$rightsbased[f_b_itq_updated$rightsbased == "0"]<- "No ITQ"
f_b_itq_updated$rightsbased[f_b_itq_updated$rightsbased == "1"]<- "ITQ"

F_B_graph_updated <- ggplot(data = f_b_itq_updated, aes( x=BvBmsy, y=FvFmsy, colour= rightsbased, size = Catch ))+
  geom_point()+
  labs(x = "B/Bmsy", y= "F/Fsmy") +
  theme_minimal()+
  theme(legend.title=element_blank())+
  ylim(-1, 5)+
  xlim(-.3, 3)+
  geom_hline(aes(yintercept=1))+
  geom_vline(aes(xintercept=1))+
  ggtitle("Fisheries 1950-2016")

#F_B_graph
ggplotly(F_B_graph_updated)
```

##From upside: the catch in ITQ versus non ITQ for most recent year

###Total Catch ITQ, No ITQ, TURF in 2012
```{r}
ITQ_projection <- readRDS("data/ITQ_projection.rds")

f_b_itq_countries <- ITQ_projection %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch", "Year", "Country", "IdOrig") %>%
  filter( itq != "NA" & iq != "NA" & ivq != "NA" & turf != "NA") %>%
  mutate(rightsbased = case_when(
    turf == TRUE ~ "2",
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

countries_itq_2012 <- f_b_itq_countries %>%
  select (Country, rightsbased, Year, IdOrig, Catch) %>%
  filter(Year == "2012")

itq_2012_rightsbasedTRUE <- filter(countries_itq_2012, rightsbased == "1")

itq_2012_rightsbasedFALSE <- filter(countries_itq_2012, rightsbased == "0")

itq_2012_rightsbasedTURF <- filter(countries_itq_2012, rightsbased == "2")

sum(itq_2012_rightsbasedTRUE$Catch, na.rm = TRUE)
#7,420,003

sum(itq_2012_rightsbasedFALSE$Catch, na.rm = TRUE)
#142,231

sum(itq_2012_rightsbasedTURF$Catch, na.rm = TRUE)
#1,301,000
```
2012 total catch:
ITQs: 7,420,003
No ITQ: 142,231 (all USA)
TURF:1,301,000 (all Japan)
**note: what units is the catch in the upside/rams data?**

###Total Catch ITQ, No ITQ, TURF in 2016
```{r}

ITQ_projection <- readRDS("data/ITQ_projection.rds")

f_b_itq_countries <- ITQ_projection %>%
  select("BvBmsy", "FvFmsy", "itq", "iq", "ivq", "turf", "Catch", "Year", "Country", "IdOrig") %>%
  filter( itq != "NA" & iq != "NA" & ivq != "NA" & turf != "NA") %>%
  mutate(rightsbased = case_when(
    turf == TRUE ~ "2",
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

countries_itq_2012_2 <- f_b_itq_countries %>%
  select (Country, rightsbased, Year, IdOrig) %>%
  filter(Year == "2012")  %>%
  select(Country, rightsbased, IdOrig)

ITQ_projection_2016 <- ITQ_projection %>%
  filter(Year == "2016") %>%
  select(IdOrig, Country, Year, Catch)

itq_2016_updated <- join(ITQ_projection_2016, countries_itq_2012_2, by= c("IdOrig", "Country"), type= "full")

itq_2016_rightsbasedTRUE <- filter(itq_2016_updated, rightsbased == "1")

itq_2016_rightsbasedFALSE <- filter(itq_2016_updated, rightsbased == "0")

itq_2016_rightsbasedTURF <- filter(itq_2016_updated, rightsbased == "2")

sum(itq_2016_rightsbasedTRUE$Catch, na.rm = TRUE)
#26,293,606

sum(itq_2016_rightsbasedFALSE$Catch, na.rm = TRUE)
#1,154,957

sum(itq_2016_rightsbasedTURF$Catch, na.rm = TRUE)
#382,370.4
```

2016 total catch:
ITQs: 27,830,933
No ITQ: 266,328,075 (all in the USA)
TURF:1,505,000 (all in Japan)

**note: This analyze does not include TURF information from DiscoverTurf GP because that data is not fishery specific enough to merge with the upside/Corbette data**

##Probability of ITQ

```{r}

species_rightsbased <- read_csv("data/species_rightsbased.csv")

gdp_all <- read_csv("/GitHub/global_fisheries_managment/data/gpd.csv")

gdp <- filter(gdp_all, current_gdp != "NA")

species_rightsbased_mutate <- species_rightsbased %>%
  mutate(i_right = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0")) %>%
  mutate(TURF = case_when(
    turf == TRUE ~ "1", 
    turf = FALSE ~"0"
  ))

##why does this value have NAs in the first place? when mutate is does line 1 first and not registering the false = 0 one. but i guess it doesn't matter beacuse its either one or the other? this is not happening for i_right so thats weird

species_rightsbased_mutate$TURF[is.na(species_rightsbased_mutate$TURF)] <- "0"

gdp <- filter(gdp_all, current_gdp != "NA")

join_gdp_rightsbased <- join(gdp, species_rightsbased_mutate, by = c("Country"), type = "full")

gdp_rightsbased <- filter(join_gdp_rightsbased, SpeciesCat != "NA" )

gdp_rightsbased$SpeciesCat <- factor(gdp_rightsbased$SpeciesCat)
gdp_rightsbased$i_right <- factor(gdp_rightsbased$i_right)

itq_glm <- glm(formula = i_right ~ current_gdp + SpeciesCat, family = "binomial", data = gdp_rightsbased)
itq_glm

summary(itq_glm)

```

### this code chunk and the one above should be the same but they are not. figure out why sinc eth two data frames are pulled from the same original info
```{r}

#ISSCAAP_species <- read_csv("data/ISSCAAP_species.csv")
#ITQ_projection <- readRDS("data/ITQ_projection.rds")

#ITQ_projection_isscaap_1 <- join(ISSCAAP_species,ITQ_projection, by = c("SciName", "SpeciesCat"), type = "full" )

#ITQ_projection_isscaap <- ITQ_projection_isscaap_1 %>%
 # filter(Country != "NA")

species_rightsbased <- read_csv("data/species_rightsbased.csv")

gdp_all <- read_csv("/GitHub/global_fisheries_managment/data/gpd.csv")

gdp <- filter(gdp_all, current_gdp != "NA")

species_rightsbased <- ITQ_projection_isscaap %>%
  select("SciName", "Country", "SpeciesCat", "itq", "iq", "ivq", "turf", "Catch") %>%
  filter( itq != "NA", iq != "NA", ivq != "NA") %>%
  mutate(i_right = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0")) %>%
  mutate(turf = case_when(
    turf == TRUE ~ "1", 
    turf = FALSE ~"0"
  )) %>%
  select("SciName", "Country", "SpeciesCat", "i_right", "turf", "Catch")

species_rightsbased$turf[is.na(species_rightsbased$turf)] <- "0"

join_gdp_rightsbased <- join(gdp, species_rightsbased, by = c("Country"), type = "full")

gdp_rightsbased <- filter(join_gdp_rightsbased, SpeciesCat != "NA" )

gdp_rightsbased$SpeciesCat <- factor(gdp_rightsbased$SpeciesCat)
gdp_rightsbased$i_right <- factor(gdp_rightsbased$i_right)

itq_glm <- glm(formula = i_right ~ current_gdp + SpeciesCat, family = "binomial", data = gdp_rightsbased)
itq_glm

summary(itq_glm)

```

##Run Turf versus ITQ probably: prob(ITQ). Turf = 1 and ITQ = 0 with the data set that was the turf/itq only one

```{r}
species_rightsbased <- read_csv("data/species_rightsbased.csv")

gdp_all <- read_csv("/GitHub/global_fisheries_managment/data/gpd.csv")

gdp <- filter(gdp_all, current_gdp != "NA")

species_rightsbased_mutate_2 <- species_rightsbased %>%
  mutate(rightbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "0",
    turf == TRUE ~ "1"))

gdp <- filter(gdp_all, current_gdp != "NA")

join_gdp_rightsbased_2 <- join(gdp, species_rightsbased_mutate_2, by = c("Country"), type = "full")

gdp_rightsbased_2 <- filter(join_gdp_rightsbased_2, SpeciesCat != "NA" )

gdp_rightsbased_2$SpeciesCat <- factor(gdp_rightsbased_2$SpeciesCat)
gdp_rightsbased_2$rightbased <- factor(gdp_rightsbased_2$rightbased)

rightbased_glm <- glm(formula = rightbased ~ current_gdp + SpeciesCat, family = "binomial", data = gdp_rightsbased_2)
rightbased_glm

summary(rightbased_glm)

```
##5.	For MPA filter only for those that are “All” for No Take
a.	Total area for each country
b.	Top ten countries by area
i.	make a bar graph and each x-axis label is a country

PA_DEF: Allowed values: 1 (meets IUCN and/or CBD PA definition); 0 (does not meet IUCN and/or CBD PA definition (currently stored outside WDPA)).

MARINE: Allowed values: 0 (100% Terrestrial PA), 1 (Coastal: marine and terrestrial PA), and 2 (100 % marine PA).

```{r}

mpa_updated <- read_csv("data/mpa_updated.csv")

mpa <- mpa_updated %>%
  select(Country, REP_M_AREA, MARINE, NO_TAKE) %>%
  filter(MARINE == "2" | MARINE == "1") %>%
  filter(NO_TAKE == "All") %>%
  group_by(Country)

mpa_area <- aggregate(mpa$REP_M_AREA, by=list(Country=mpa$Country), FUN=sum)
mpa_area

names(mpa_area) <-c("Country", "Area")

mpa_area_10 <- top_n(mpa_area, 10, Area)
mpa_area_10

ggplot(mpa_area_10, aes(x = Country, y = Area)) +
  geom_bar(stat = "identity")

mpa_area_nojurisdiction <-filter(mpa_area, Country != "Areas Beyond National Jurisdiction")
mpa_area_nojurisdiction_10 <- top_n(mpa_area_nojurisdiction, 10, Area)
mpa_area_nojurisdiction_10

ggplot(mpa_area_nojurisdiction_10, aes(x = Country, y = Area)) +
  geom_bar(stat = "identity")

```