### 2018-12-18 draft_figures_12.18.18.Rmd
estimates of total catch for turf, itq, no itq

---
title: "draft_figures_12.18.18"
author: "Juliette Verstaen"
date: "12/18/2018"
output: pdf_document
toc: true
toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE }

suppressPackageStartupMessages({
  library(tidyverse)
  library(tidyr)
  library(here)
  library(ggplot2)
  library(dplyr)
  library(readxl)
  library(devtools)
  library(plotly)
  library(base)
  library(rio)
  library(plyr)
})
```

##KOBE Plots with only most recent year of fisheries
###RAM only
No data: ITQ = FALSE

```{r}
fisheries_recent <- read_csv("data/fisheries_recent.csv")

fisheries_recent$itq[is.na(fisheries_recent$itq)] <- "FALSE"
fisheries_recent$ivq[is.na(fisheries_recent$ivq)] <- "FALSE"
fisheries_recent$iq[is.na(fisheries_recent$iq)] <- "FALSE"
fisheries_recent$turf[is.na(fisheries_recent$turf)] <- "FALSE"
```

```{r}

fisheries_KOBE_ram <- fisheries_recent %>%
  filter(Dbase == "RAM") %>%
  mutate(rightsbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

fisheries_KOBE_ram$rightsbased[fisheries_KOBE_ram$rightsbased == "0"]<- "No ITQ"
fisheries_KOBE_ram$rightsbased[fisheries_KOBE_ram$rightsbased == "1"]<- "ITQ"

 ggplot(data = fisheries_KOBE_ram, aes( x=BvBmsy, y=FvFmsy, colour= rightsbased, size = Catch ))+
  geom_point()+
  labs(x = "B/Bmsy", y= "F/Fsmy") +
  theme_minimal()+
  theme(legend.title=element_blank())+
  ylim(0, 2.5)+
  xlim(0, 2.5)+
  geom_hline(aes(yintercept=1))+
  geom_vline(aes(xintercept=1))+
  ggtitle("KOBE Plot: RAM only Fisheries")

```

##KOBE Plots with only most recent year of fisheries
###All Data Sources
No data: ITQ = FALSE
```{r}

fisheries_KOBE <- fisheries_recent %>%
  mutate(rightsbased = case_when(
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0"))

fisheries_KOBE$rightsbased[fisheries_KOBE$rightsbased == "0"]<- "No ITQ"
fisheries_KOBE$rightsbased[fisheries_KOBE$rightsbased == "1"]<- "ITQ"

ggplot(data = fisheries_KOBE, aes( x=BvBmsy, y=FvFmsy, colour= rightsbased, size = Catch ))+
  geom_point()+
  labs(x = "B/Bmsy", y= "F/Fsmy") +
  theme_minimal()+
  theme(legend.title=element_blank())+
  ylim(0, 2.5)+
  xlim(0, 2.5)+
  geom_hline(aes(yintercept=1))+
  geom_vline(aes(xintercept=1))+
  ggtitle("KOBE Plot:  All Fisheries")

```

##Total Catches in most recent year of ITQ, no ITQ, and Turfs
```{r}

fisheries_rightsbased <- fisheries_recent %>%
  mutate(rightsbased = case_when(
    turf == TRUE ~ "2",
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0", 
    ))

turfs <- filter(fisheries_rightsbased, rightsbased == "2")
itq <- filter(fisheries_rightsbased, rightsbased == "1")
no_itq <- filter(fisheries_rightsbased, rightsbased == "0")

sum(turfs$Catch, na.rm = TRUE)
#20,217,168

sum(itq$Catch, na.rm = TRUE)
#1,154,957

sum(no_itq$Catch, na.rm = TRUE)
#382,370.4
```
turf = 0
itq = 4,056,613
no itq = 66,156,579

## adding turf information to upside projection
```{r}

turf_itq_isscaap <- read_csv("data/turf_itq_isscaap.csv")

turf_only <- turf_itq_isscaap %>%
  select(SciName, Country, turf)

fisheries_recent_generousturf_1 <- select(fisheries_recent, Country, assess_id_short, Year.x, CommName, Biomass, Catch, BvBmsy, FvFmsy, Dbase, SciName, IdLevel, SpeciesCat.x, Profits, MSY, Price, g, k, c, phi, itq, ivq, iq)

fisheries_recent_generousturf <- merge(turf_only, fisheries_recent_generousturf_1, by = c("Country", "SciName"), all = "TRUE")

fisheries_recent_generousturf$turf[is.na(fisheries_recent_generousturf$turf)] <- "FALSE"
fisheries_recent_generousturf$itq[is.na(fisheries_recent_generousturf$itq)] <- "FALSE"
fisheries_recent_generousturf$ivq[is.na(fisheries_recent_generousturf$ivq)] <- "FALSE"
fisheries_recent_generousturf$iq[is.na(fisheries_recent_generousturf$iq)] <- "FALSE"

fisheries_recent_generousturf_rightsbased <- fisheries_recent_generousturf %>%
  mutate(rightsbased = case_when(
    turf == TRUE ~ "2",
    itq == FALSE & iq == FALSE & ivq == FALSE ~ "0",
    itq == TRUE | iq == TRUE | ivq == TRUE ~ "1",
    ))

turfs <- filter(fisheries_recent_generousturf_rightsbased, rightsbased == "2")
itq <- filter(fisheries_recent_generousturf_rightsbased, rightsbased == "1")
no_itq <- filter(fisheries_recent_generousturf_rightsbased, rightsbased == "0")

sum(turfs$Catch, na.rm = TRUE)

sum(itq$Catch, na.rm = TRUE)

sum(no_itq$Catch, na.rm = TRUE)

69025951/78938573

```

Generous Turf Catch Estimates:

turf = 1,357,098 -> 1.72% global catch
itq= 8,555,524 -> 10.8% global catch
no itq =  69,025,951 -> 87.44% global catch

Total: 78,938,573