### 2019-01-10 documents/dan-figs-and-calcs.Rmd
a few random things

---
title: "dan_global_fisheries"
output: html_notebook
---

```{r}
library(tidyverse)
library(readxl)
library(tidyxl)
library(here)

# source(here("scripts","2016_projection_updated.R"))

fisheries_recent <- read_csv(here("processed_data","fisheries_recent.csv"))%>% 
  janitor::clean_names()

```

```{r}

```

# rbfm

However, while these fisheries make up XX\% of global captures
```{r}

global_summaries <- fisheries_recent %>% 
  group_by(dbase) %>% 
  summarise(total_catch = sum(catch, na.rm = T),
            n = n_distinct(assess_id_short))

```

# mpas

```{r}

mpas <- readxl::read_xlsx(here("data","MPA Analysis. Country Analysis.xlsx"), sheet = "clean_mpa") %>% janitor::clean_names()

```